---
title: "Buprenorphine Q1 2010 - Q1 2020"
author: "Michael Maguire, MS"
date: "3/18/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries

```{r libload, message = FALSE}
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(purrr)
library(readr)
library(readxl)
library(stringr)
library(tidylog)
library(viridis)
```

### Step 1: Read in SK's data set with drug names.

```{r skread, include = TRUE}
drugs <-
  read_xlsx(
    path = "./data/raw/Buprenorphine drug names_SK.xlsx",
    .name_repair = "universal"
  )

drugs_vec <-
  drugs$Drug.name

drugs_vec
```

### Step 2: Get the states that SK requested.

```{r skstates, include = TRUE}
sk_states <- 
  data.frame(
    state = c(
      "California", "Delaware", "Georgia", "Illinois", "Florida", "Idaho", "Indiana", "Kansas", "Louisiana",
      "Mississippi", "Montana", "Nevada", "North Dakota", "South Dakota", "Wyoming", "Tennessee"
    )
  )

states <-
  bind_cols(
    state.name,
    tolower(state.abb)
  ) %>%
  rename(state = `...1`, abb = `...2`)

states

states_req <-
  states %>%
    inner_join(sk_states, by = "state") %>%
  select(abb)

state_vec <-
  as_vector(states_req)

state_vec

```

### Step 3: Pull in SDUD dataset.

Note: location of data set is a personal directory, so I have masked it.

```{r locload, include = FALSE}
loc <- "C:\\Users\\michaelqmaguire2\\Dropbox (UFL)\\01_projects\\us-medicaid-drug-utilization\\us-medicaid-drug-utilziation\\data\\clean\\medicaid-sdud-2000-2020.rds"
```

```{r sdudload, echo = TRUE}
sdud_2010_2020 <-
  read_rds(
    file = loc
  ) %>%
  filter(state %in% state_vec & (year %in% c(2010:2019) | year == 2020 & quarter == 1))
```

Examine years and quarters to ensure the correct period is pulled.

```{r yrqtrcheck, echo = TRUE}
sdud_2010_2020 %>%
  distinct(year, quarter) %>%
  arrange(year, desc(quarter))
```

### Step 4: Flag records containing generic name.

```{r genflag, echo = TRUE}
sdud_2010_2020_gennmes <- 
  sdud_2010_2020 %>%
  mutate(
    buprenorphine_flag = case_when(
      str_detect(string = gennme_c, pattern = regex("buprenorphine", ignore_case = TRUE)) ~ "1",
      TRUE ~ "0"
    )
  )
```

Checking flags.

```{r flagcheck, echo = TRUE}
sdud_2010_2020_gennmes %>%
  select(buprenorphine_flag) %>%
  map(., janitor::tabyl)
```

### Step 5: Aggregate by state, year, quarter and sum.

```{r aggin, echo = TRUE}
# Filter off on flags created above.
sdud_2010_2020_bup_flags <-
  sdud_2010_2020_gennmes %>%
  filter(
    buprenorphine_flag == "1"
  )
# Create data set containing generic name, brand name, and ndc number.
ndc_gen_brand_names <- 
  sdud_2010_2020_bup_flags %>%
  distinct(gennme_c, prodnme, ndc)
# Examine data set.
ndc_gen_brand_names
# Create aggregate by state, year, quarter, and suppression.
sdud_2010_2020_bup_rx <-
  sdud_2010_2020_bup_flags %>%
  group_by(state, year, quarter, gennme_c, suppression_used) %>%
  summarize(total_prescriptions = sum(number_of_prescriptions))
# Create aggregate by state, year, quarter, generic name, and suppression. 
sdud_2010_2020_bup <- 
  sdud_2010_2020_bup_flags %>%
  group_by(state, year, quarter, suppression_used) %>%
  summarize(total_prescriptions = sum(number_of_prescriptions))
```

### Step 6: Plot number of prescriptions over time.

```{r plots, echo = TRUE, fig.width = 12, fig.height = 8}
# Plot showing number of prescriptions by year and quarter.
ggplot(data = sdud_2010_2020_bup) +
  geom_col(aes(x = paste0(year, "-", quarter), y = total_prescriptions), fill = "forestgreen", alpha = 0.95) +
  scale_y_continuous(labels = scales::comma) + 
  theme_ipsum_rc(axis_title_just = "ct") +
  ggtitle("Number of Prescriptions by Year and Quarter") +
  xlab("Year-Quarter") +
  ylab("Total Number of Prescriptions") +
  theme(
    axis.text.x = element_text(color = "black", angle = 45, vjust = 0.5, hjust = 0.75),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black", size = 10, vjust = -4),
    axis.title.y = element_text(color = "black", size = 10),
  ) +
  coord_cartesian(expand = FALSE)
# Plot showing number of prescriptions by year, quarter, and generic name.
ggplot(data = sdud_2010_2020_bup) +
  geom_col(aes(x = paste0(year, "-", quarter), y = total_prescriptions), fill = "forestgreen", alpha = 0.95) +
  scale_y_continuous(labels = scales::comma) + 
  theme_ipsum_rc(axis_title_just = "ct") +
  ggtitle("Number of Prescriptions by Year and Quarter") +
  xlab("Year-Quarter") +
  ylab("Total Number of Prescriptions") +
  theme(
    axis.text.x = element_text(color = "black", angle = 45, vjust = 0.5, hjust = 0.75),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black", size = 10, vjust = -4),
    axis.title.y = element_text(color = "black", size = 10),
  ) +
  coord_cartesian(expand = FALSE)
```
```{r plotrx, fig.height = 12, fig.width = 16}
# Plot showing number of prescriptions by year, quarter, and state.
ggplot(data = sdud_2010_2020_bup) +
  geom_col(aes(x = paste0(year, "-", quarter), y = total_prescriptions, fill = state), alpha = 0.95) +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = scales::comma) + 
  theme_ipsum_rc(axis_title_just = "ct") +
  ggtitle("Number of Prescriptions by Year, Quarter, and State") +
  xlab("Year-Quarter") +
  ylab("Total Number of Prescriptions") +
  theme(
    axis.text.x = element_text(color = "black", angle = 90, size = 8, hjust = 0.25, vjust = 0.25),
    axis.text.y = element_text(color = "black"),
    axis.title.x = element_text(color = "black", size = 10),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "none"
  ) +
  facet_wrap(~state) +
  coord_cartesian(expand = FALSE)
```