---
title: "Buprenoprhine 2013Q4 - 2020Q1"
author: "Michael Quinn Maguire, MS"
date: "2/9/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages Used

```{r packageLoad, message = FALSE}
library(data.table)
library(hrbrthemes)
library(tidyverse)
library(tidylog)
```

## Create 'state' data frame and vectorize the results.

```{r stateCreation}
skStates <- setNames(
  as.data.table(
    c('Alabama', 'Florida', 'Idaho', 'Indiana', 'Kansas', 'Louisiana', 'Mississippi', 'Montana',
      'Nebraska', 'Nevada', 'North Carolina', 'North Dakota', 'South Dakota', 'Tennessee', 'Wyoming'
      )
    ),
  "state")

statesAbb <- setNames(as.data.table(cbind(state.name, state.abb)), c("state", "abbreviation"))

skFinal <-
  statesAbb[skStates, on = .(state = state)]

statesOfInterest <- skFinal$abbreviation

skFinal
```

## Read in SDUD file 

> Note: SDUD file is masked.

```{r sdudLoad}
sdud <- fread(
  file_location,
  colClasses = c("proper_ndc" = "character")
)

sdud
```

## Create an initial aggregation without quarters.

```{r initialAgg}
yrSt <-
  sdud[
    i = year %in% c(2013:2020) & state %in% statesOfInterest & str_detect(gennme, pattern = "Buprenorphine"),
    j = .(utilization.type, strngth, mstfmds, year, proper_ndc, quarter, state, suppression, numberrx, prodnme, gennme)
  ]

setorder(yrSt, state, year, quarter, proper_ndc)

yrSt
```

## Check brand names.

```{r brandNameCheck}
unique(yrSt$prodnme, by = 'prodnme')
```

## Create variables representing whether row is in requested date range or not.

```{r inRange}
yrStRange <- yrSt[
    i = ,
    j = `:=` (inRange = ifelse((year == 2013 & quarter < 4) | (year == 2020 & quarter > 1), 0, 1), yearqtr = paste0(year, '-', quarter))
  ]

yrStRange
```

## Running some checks to ensure that correct dates are being pulled

```{r checks}
yrStRange |>
  janitor::tabyl(inRange)

yrStRange |>
  filter(inRange == 1) |>
  distinct(yearqtr)

yrStRange |>
  filter(inRange == 0) |>
  distinct(yearqtr)
```

## Create the aggregated data set

```{r finalAggregate}
drugsAggStateProdnmeRange <- yrStRange[
    i = inRange == 1,
    j  = .(totalRX = sum(numberrx)),
    by = c("utilization.type", "year", "state", "quarter", "gennme", "prodnme", "strngth", "mstfmds", "suppression")
  ]

drugsAggStateProdnmeRange
```

## Plot total prescriptions by year, quarter, and generic name.

```{r yearQtrGenericAgg, fig.width = 20, fig.height = 12}
drugsAggStateProdnmeRange |> filter(suppression == 'F') |>
ggplot() +
  geom_col(
    aes(
      x = as.factor(paste0(year, '-', quarter)),
      y = totalRX,
      fill = gennme
    )
  ) +
  theme_ipsum_rc() +
  theme(
    axis.text.x = element_text(angle = 90)
  ) + 
  scale_y_continuous(labels = scales::comma) +
  scale_fill_viridis_d(direction = -1)
```

## Plot total prescriptions by year, quarter, and brand name.

```{r yearQtrProdAgg, fig.width = 20, fig.height = 12}
drugsAggStateProdnmeRange |> filter(suppression == 'F') |>
ggplot() +
  geom_col(
    aes(
      x = as.factor(paste0(year, '-', quarter)),
      y = totalRX,
      fill = prodnme
    )
  ) +
  theme_ipsum_rc() +
  theme(
    axis.text.x = element_text(angle = 90)
  ) + 
  scale_y_continuous(labels = scales::comma) +
  scale_fill_viridis_d(direction = -1)
```

## Plot total prescriptions by year, quarter, generic name, and state.

```{r yearQtrStateGeneric, fig.width = 20, fig.height = 12}
drugsAggStateProdnmeRange |> filter(suppression == 'F') |>
  ggplot() +
  geom_col(
    aes(
      x = as.factor(paste0(year, '-', quarter)),
      y = totalRX,
      fill = gennme
    )
  ) + 
  facet_wrap(~state) +
  theme_ipsum_rc() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_viridis_d(direction = -1)
```

## Plot total prescriptions by year, quarter, brand name, and state.

```{r yearQtrStateBrand, fig.width = 20, fig.height = 12}
drugsAggStateProdnmeRange |> filter(suppression == 'F') |>
  ggplot() +
  geom_col(
    aes(
      x = as.factor(paste0(year, '-', quarter)),
      y = totalRX,
      fill = prodnme
    )
  ) + 
  facet_wrap(~state) +
  theme_ipsum_rc() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_viridis_d(direction = -1)
```

